# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

First we unzip the activity.zip in this repository, then open and load the csv file

``` {r echo=TRUE}
unzip(zipfile = "activity.zip", overwrite = TRUE)
table <- read.csv("./activity.csv")
table$date <- as.Date(table$date)
table$steps <- as.numeric(as.character(table$steps))
table$interval <- as.numeric(as.character(table$interval))
head(table)
```


Preprocess the data

``` {r echo=TRUE}
steps_date <- aggregate(steps ~ date, data = table, FUN = sum)
str(steps_date)
```


``` {r echo=TRUE}
head(steps_date)
```


## What is mean total number of steps taken per day?

Plot a histogram that plot the steps per day.
``` {r echo=TRUE}
barplot(steps_date$steps, names.arg = steps_date$date, xlab = "Date", ylab = "Steps")
```

Mean and median are 10766.19 and 10765:

``` {r echo=TRUE}
steps_mean <- mean(steps_date$steps)
steps_mean
```


``` {r echo=TRUE}
steps_median <- median(steps_date$steps)
steps_median
```


## What is the average daily activity pattern?

Plot a time series of the 5-minute interval and the average number of steps taken, averaged across all days


``` {r echo=TRUE}
steps_per_day <- aggregate(steps ~ interval, data = table, FUN = mean)
plot(steps_per_day, type = "l", xlab = "Interval", ylab = "Steps")
```


Find the maximum number of steps:


``` {r echo=TRUE}
steps_per_day$interval[which.max(steps_per_day$steps)]
```



## Imputing missing values

How many missing values in the dataset:


``` {r echo=TRUE}
sum(is.na(table))
```


``` {r echo=TRUE}
table <- merge(table, steps_per_day, by = "interval", suffixes = c("", ".y"))
na <- is.na(table$steps)
table$steps[na] <- table$steps.y[na]
table <- table[, c(1:3)]
```



compare to the previous figure, the median has changed a little and is now exactly equal to the mean
 
 
``` {r echo=TRUE}
steps_date <- aggregate(steps ~ date, data = table, FUN = sum)
barplot(steps_date$steps, names.arg = steps_date$date, xlab = "Date", ylab = "Steps")
```


``` {r echo=TRUE}
mean(steps_date$steps)
```

``` {r echo=TRUE}
median(steps_date$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

Find days are either weekdays or weekends.

``` {r echo=TRUE}
table$weekdayorend <- ifelse(weekdays(table$date) == "Saturday" | weekdays(table$date) == "Sunday", "weekend", "weekday")
head(table)
```


Plot weekend and weekday figures to see the differences:

``` {r echo=TRUE}
par(mfrow = c(2, 1))
steps_type <- aggregate(steps ~ interval, data = table, subset = table$weekdayorend == "weekend", FUN = mean)
plot(steps_type, main = "weekend", type = "l", xlab = "Interval", ylab = "Steps")
steps_type <- aggregate(steps ~ interval, data = table, subset = table$weekdayorend == "weekday", FUN = mean)
plot(steps_type, main = "weekday", type = "l", xlab = "Interval", ylab = "Steps")
```

Yes there are differences, for example on weekday it has higher peak.
